# File Description 

This notebook is to compile all pre-processing code needed for display in my written report. I will be laying out all relative code in organized cells before converting them into standalone functions in a document called `preprocessing_functions.R` located in the `scripts` folder. I will also be re-factoring the original code to better adhere to styling standards as laid out by [Hadley Wickham](http://adv-r.had.co.nz/Style.html).

## Load preliminary data
```{r}
library(tidyverse)
temp <- list.files(path = "../data/raw-data/")

macro_data <- read.csv(file = paste("../data/raw-data/", temp[1], sep = ""))
pop_growth <- read.csv("../data/raw-data/Population-Growth.csv") # Yearly so won't include

#Left join all other data (except growth) onto macroData
for (i in temp[c(2:6,8)]){
  path <- paste("../data/raw-data/", i, sep = "")
  temp_df <- read.csv(file = path)
  macro_data <- merge(macro_data, temp_df, by = "DATE", all.x = TRUE)
}

#Convert "DATE" column to Date type
macro_data$DATE <- as.Date(macro_data$DATE)

post_2003 <- macro_data %>%
  filter(DATE >= "2003-01-01")
```

## Test preprocessing function
```{r}
source("../scripts/preprocessing_functions.R")
preliminary_data <- load_preliminary_data()
```

## Display preliminary data as a table
```{r}
library(flextable)
library(skimr)

post_2003 %>%
  skimr:: skim() %>%
  dplyr::select(skim_type, skim_variable, n_missing, numeric.hist) %>%
  flextable::flextable()%>%
  set_caption(
    caption = as_paragraph(
      as_chunk("Table 1: Structure of Intermediate Data",
               props = fp_text_default(font.family = "Times New Roman",
                                       font.size = 12,
                                       italic = TRUE)))) %>%
  set_header_labels(values = list(
    skim_type = "Type",
    skim_variable = "Name",
    n_missing = "NA values",
    numeric.hist = "Histogram")) %>%
  set_table_properties(layout = "autofit")%>%
  bg(i = seq(2,ncol(post_2003),2), bg = "light grey")%>%
  font(part = "all", fontname = "Times New Roman") %>%
  fontsize(size = 12) %>%
  border(i = c(1:ncol(post_2003)), border.top = fp_border_default(color = "black"))
```

## Test preprocessing function
```{r}
source("../scripts/preprocessing_functions.R")
display_preliminary_table(post_2003)
```

## Graph raw GDP
```{r}
library(ggthemes)
#Graph raw GDP
min_date <- as.Date(min(post_2003$DATE))
max_date <- as.Date(max(post_2003$DATE))
date_limit <- c(min_date, max_date)

post_2003 %>%
  ggplot(aes(x = DATE, y = GDP.x)) +
  geom_point(na.rm = T, cex = 1) + 
  xlab("") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_x_date(limit = date_limit)
```

## Graph imputed GDP
```{r}
#Extract GDP
GDP <- post_2003$GDP.x

palette <- c("#5F0F40",
             "#9A031E",
             "#FB8B24",
             "#E36414",
             "#0F4C5C")

#Impute NA values using trend
impute_GDP <- imputeTS::na_seadec(GDP)
post_2003$impute_GDP <- impute_GDP

#Create imputed category
post_2003$imputed <- as.factor(
  ifelse(
    is.na(
      ifelse(post_2003$GDP.x ==post_2003$impute_GDP, 0, 1)
      ) 
    == T, 1,0))


#Graph imputed GDP
post_2003 %>%
  ggplot(aes(x = DATE, y = impute_GDP, color = imputed)) + 
  geom_point(na.rm = T, cex = 1, alpha = 0.50) + 
  geom_line(aes(x = DATE, y = impute_GDP), na.rm = T, alpha = 1) + 
  xlab("") +
  theme_bw() + 
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(limit= date_limit) + 
  scale_color_manual(values = c(palette[1], palette[3]))
```

## Test preprocessing function
```{r}
#Graph pre impute GDP
source("../scripts/preprocessing_functions.R")

plot_pre_imputed_GDP(post_2003)
```

## Test preprocessing function
```{r}
#Impute GDP
source("../scripts/preprocessing_functions.R")
post_2003 <- load_preliminary_data()
post_2003$imputed_GDP <- impute_GDP(post_2003)
post_2003 <- create_imputed_column(post_2003, "GDP.x", "imputed_GDP")
```

## Test preprocessing function
```{r}
#Graph impute GDP
source("../scripts/preprocessing_functions.R")
post_2003 <- load_preliminary_data()
post_2003$imputed_GDP <- impute_GDP(post_2003)
post_2003 <- create_imputed_column(post_2003, "GDP.x", "imputed_GDP")

plot_imputed_GDP(post_2003, date_col = "DATE", imputed_GDP_col = "imputed_GDP",
                 imputed_cat_col = "imputed")
```

## Remove low quality columns
```{r}
# Get rid of the following columns (with reasons):
  #MORTGAGE3OUS seems to be low quality -- remove
  #Unsure about GDP.y's sources-- remove
  #interest_rate is a duplicate of FEDFUNDS-- remove
  #home_price_index is a duplicate of CSUSHPISA -- remove
  #unemployment_rate is a duplicate of UNRATE -- remove
  #unsure of data quality of house_for_sale_or_sold -- remove

#Duplicate post_2003 to keep as intermediate data
post_2003_viz <- post_2003

#List of columns to delete
columns_to_delete <- c("MORTGAGE30US", "GDP.y", "interest_rate",
                       "home_price_index", "unemployment_rate",
                       "house_for_sale_or_sold")

#Loop through and set to NULL
for (i in columns_to_delete){
  post_2003_viz2[,i] <- NULL
}
```

No need for a separate function for this process ^

## Remove GDP and imputed columns for modeling
```{r}
post_2003_model <- post_2003_viz

#Remove GDP.x from post2003 and only leave imputeGDP, remove imputeYN
post_2003_model$GDP.x <- NULL
post_2003_model$imputed <- NULL

```

No need for a separate function for this process ^

## Rename columns
```{r}

test <- post_2003_viz %>% select(-c(GDP.x, imputed))

colnames(test)

new_names = c("date", "urban_cpi", "fed_funds_rt", "hpi", "build_permits",
              "const_price_idx", "delinq_rt", "house_subsidies", "income",
              "mortgage_rt", "const_unit", "tot_house", "tot_const_spend",
              "urban_pop", "unem_rt", "imputed_gdp")

colnames(test) <- new_names

#Rename columns
post_2003_model <- post_2003_model %>% 
  rename(date = DATE,
         urban_cpi = CPIAUCSL,
         fed_funds_rt = FEDFUNDS,
         hpi = CSUSHPISA,
         build_permits = building_permits,
         const_price_idx = const_price_index,
         delinq_rt = delinquency_rate, 
         house_subsidies = housing_subsidies,
         income = income,
         mortgage_rt = mortgage_rate,
         const_unit = construction_unit,
         tot_house = total_houses,
         tot_const_spend = total_const_spending,
         urban_pop = urban_population,
         unem_rate = UNRATE,
         imputed_gdp = imputed_GDP
         )
```

Function not needed ^

# Check and display NA values
```{r}
#NA check -- still some NA values remaining
colSums(is.na(post_2003_model))

na_vals <- tibble:: rownames_to_column(data.frame(na_values = colSums(is.na(post_2003_model))), "variable")

#Display table of NA values
na_vals %>%
  flextable:: flextable() %>%
  set_caption(
    caption = as_paragraph(
      as_chunk("Table 2: Remaining NA Values After Processing",
               props = fp_text_default(font.family = "Times New Roman",
                                       font.size = 12,
                                       italic = TRUE)))) %>%
  set_header_labels(values = list(
    variable = "Variable",
    na_values = "Number of Missing Values")) %>%
  set_table_properties(layout = "autofit") %>%
  bg(i = seq(2, nrow(na_vals), 2), bg = "light grey")%>%
  font(part = "all", fontname = "Times New Roman") %>%
  fontsize(size = 12) %>%
  border(i = c(1:nrow(na_vals)), border.top = fp_border_default(color = "black"))
```

## Test preprocessing function
```{r}
#Display NA values in a flextable before removing them
source("../scripts/preprocessing_functions.R")
display_intermediate_nas(post_2003_model)
```

## Remove NAs and extract month and date
```{r}
#NAs are low enough that I will just remove them
post_2003_clean <- na.omit(post_2003_model)

#Add date features -- don't need day since monthly data
post_2003_clean$year <- lubridate::year(post_2003_clean$date)
post_2003_clean$month <- lubridate::month(post_2003_clean$date)


post_2003_clean$year <- as.Date(post_2003_clean$year)
```

## Test preprocessing function
```{r}
source("../scripts/preprocessing_functions.R")
clean_data <- final_preprocessing(intermediate_data = post_2003_model)
```

