# File Description 

This notebook is to compile all visualization code needed for display in my written report. I will be laying out all relative code in organized cells before converting them into standalone functions in a document called `preprocessing_functions.R` located in the `scripts` folder. I will also be re-factoring the original code to better adhere to styling standards as laid out by [Hadley Wickham](http://adv-r.had.co.nz/Style.html).

## Globally-referenced palette
```{r}
palette <- c("#5F0F40",
             "#9A031E",
             "#FB8B24",
             "#E36414",
             "#0F4C5C")
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)
library(grid)
library(ggthemes)
library(ggcorrplot)
library(lubridate)
```

## Reading in visualization data
```{r}
# Load data
viz_data <- read.csv("../data/cleaned/visualizationData.csv") 
#Leave NAs in for GDP.x and Imputed GDP comparison

# Assert types
viz_data$DATE <- as.Date(viz_data$DATE)
viz_data$year <- lubridate::year(viz_data$DATE)
viz_data$month <- lubridate::month(viz_data$DATE)
```

## Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
v_data <- load_viz_data()
```

## Visualize GDP Imputation
```{r}
p1 <- viz_data %>%
  ggplot(aes(x=DATE, y=GDP.x)) +
  geom_point(cex = 1) + 
  geom_smooth(se = F, color = palette[3]) + 
  geom_vline(xintercept = as.Date("2020-03-15"),
             linetype = 2, linewidth = 0.50, color = palette[2]) +
  geom_vline(xintercept = as.Date("2008-09-15"),
             linetype = 2, linewidth = 0.50, color = palette[2]) + 
  annotate(x = ymd("2020-03-15"), y = +Inf, label = "COVID-19 Starts", vjust = 2, geom = "label") +
  annotate(x = ymd("2008-09-15"), y = +Inf, label = "Peak of Financial Crisis", vjust = 2, geom = "label") +
  labs(x = "", y = "GDP (Quarterly)",
       title = "US GDP", 
       subtitle = "Reported quarterly (smoothed estimate in orange)") + 
  theme_bw() +
  theme(plot.title = element_text(face = "bold", size = 25, hjust = 0),
        plot.subtitle = element_text(face = "italic", size = 10, hjust = 0))

p2 <- viz_data %>%
  ggplot(aes(x=DATE, y=imputeGDP)) +
  geom_point(cex = 1) + 
  geom_vline(xintercept = as.Date("2020-03-15"),
             linetype = 2, linewidth = 0.50, color = palette[2]) +
  geom_vline(xintercept = as.Date("2008-09-15"),
             linetype = 2, linewidth = 0.50, color = palette[2]) + 
  annotate(x = ymd("2020-03-15"), y = +Inf, label = "COVID-19 Starts", vjust = 2, geom = "label") +
  annotate(x = ymd("2008-09-15"), y = +Inf, label = "Peak of Financial Crisis", vjust = 2, geom = "label") +
  labs(x = "", y = "GDP (Monthly Imputed from Quarterly Values)",
       title = "US GDP", 
       subtitle = "Monthly (Imputed)") + 
  theme_bw() +
  theme(plot.title = element_text(face = "bold", size = 25, hjust = 0),
        plot.subtitle = element_text(face = "italic", size = 10, hjust = 0))

p3 <- gridExtra::grid.arrange(p1, p2, nrow = 1)
```

## Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
#Using vdata loaded in last test
gdp_plot <- plot_GDP_side_by_side(v_data)
```

## Plot hpi values
```{r}
p4 <- viz_data %>%
  ggplot(aes(x = DATE, y = CSUSHPISA, color = FEDFUNDS)) + 
  geom_line(aes(linewidth = FEDFUNDS), lineend = "round") + 
  geom_vline(xintercept = as.Date("2020-03-15"),
           linetype = 2, linewidth = 0.50, color = palette[2]) +
  geom_vline(xintercept = as.Date("2008-09-15"),
             linetype = 2, linewidth = 0.50, color = palette[2]) + 
  annotate(x = ymd("2020-03-15"), y = +Inf, 
           label = "COVID-19 Starts", vjust = 2, geom = "label") +
  annotate(x = ymd("2008-09-15"), y = +Inf, 
           label = "Peak of Financial Crisis", vjust = 2, geom = "label") +
  labs(x = "", y = "National Home Price Index",
       title = "S&P CoreLogic Case-Shiller U.S. National Home Price Index over Time", 
       subtitle = "Changes in Federal Funds Rate shown using color and size") + 
  theme_bw() +
  scale_color_gradient(name = "Federal Funds Rate",high = palette[2], low = palette[4]) +
  scale_linewidth(name = "Federal Funds Rate", range = c(0.5,2.5)) +
  guides(color=guide_legend(override.aes=list(fill=NA))) + 
  theme(plot.title = element_text(face = "bold", size = 25, hjust = 0),
        plot.subtitle = element_text(face = "italic", size = 10, hjust = 0),
        legend.key = element_rect(fill = "transparent"),
        legend.title = element_text(size = 10, face = "bold"),
        legend.background = element_rect(linewidth = 0, linetype = "blank"))

p4
```

## Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
#Use vdata loaded from previous test
plot_hpi_values(v_data)
```

## Line plot matrix 
```{r}
plots <- list()
ind = 1
for(i in 1: ncol(viz_data)){
  if(i == 1){
    print("pass")
  }else if(i == 4){
    plots[[ind]] <- ggplot(viz_data, aes_string(x = viz_data[,1], y = viz_data[,i])) +
      geom_point() + 
      labs(x = "", y = colnames(viz_data)[i]) +   
      geom_vline(xintercept = as.Date("2020-03-15"),
                 linetype = 2, linewidth = 0.50, color = palette[2]) +
      geom_vline(xintercept = as.Date("2008-09-15"),
                 linetype = 2, linewidth = 0.50, color = palette[2]) + 
      annotate(x = ymd("2020-03-15"), y = +Inf, 
               label = "C19", vjust = 2, geom = "label") +
      annotate(x = ymd("2008-09-15"), y = +Inf, 
               label = "Fin. Crisis", vjust = 2, geom = "label") + 
      theme_bw()
    ind = ind + 1
  }else if(i == 18){
    print("pass")
  }
  else{
    plots[[ind]] <- ggplot(viz_data, aes_string(x = viz_data[,1], y = viz_data[,i])) +
      geom_line() + 
      labs(x = "", y = colnames(viz_data)[i]) +
      geom_vline(xintercept = as.Date("2020-03-15"),
                 linetype = 2, linewidth = 0.50, color = palette[2]) +
      geom_vline(xintercept = as.Date("2008-09-15"),
                 linetype = 2, linewidth = 0.50, color = palette[2]) + 
      annotate(x = ymd("2020-03-15"), y = +Inf, 
               label = "C19", vjust = 2, geom = "label") +
      annotate(x = ymd("2008-09-15"), y = +Inf, 
               label = "Fin. Crisis", vjust = 2, geom = "label") + 
      theme_bw()
    ind = ind + 1
  }
}

p5 <- grid.arrange(grobs = plots, nrow = 4, 
                   top = textGrob(
                     substitute(
                       paste(
                         bold("Numeric Variables by Year")))))
```

##Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
#use v_data from previous test
grid <- plot_line_matrix(v_data)
```

## HPI and CPI over time
```{r, warning=FALSE}
viz_data %>%
  select(year, CSUSHPISA, CPIAUCSL) %>%
  na.omit() %>%
  group_by(year) %>%
  summarize(avgHPI = mean(CSUSHPISA),
            avgCPI = mean(CPIAUCSL)) %>%
  ggplot() +
  geom_bar(aes(x = year, y = avgHPI, 
               fill = factor(year, order = TRUE, levels = c(unique(viz_data$year)))), 
           stat = "identity") +
  scale_fill_manual(values = c("#5F0F40","#67153F","#6F1B3D",
                               "#76223C","#7E283A","#862E39",
                               "#8E3438","#963A36","#9D4135",
                               "#A54733","#AD4D32","#B55331",
                               "#BD592F","#C4602E","#CC662C",
                               "#D46C2B","#DC722A","#E47828",
                               "#EB7F27","#F38525","#FB8B24")) +
  geom_line(aes(x = year, y = avgCPI, color = year)) + 
  scale_y_continuous( name = "Average HPI", 
                      sec.axis = sec_axis(trans = ~., name = "Average CPI")) + 
  labs(x = "", y = "Average HPI", fill = "Year") + 
  scale_color_gradient(low = palette[1], high = palette[4]) + 
  guides(color = "none", fill = guide_legend(nrow = 10)) + 
  theme_bw() + 
  theme(axis.title.y= element_text(size = 7))
```

```{r}
viz_data %>%
      select(year, CSUSHPISA, CPIAUCSL) %>%
      na.omit() %>%
      group_by(year) %>%
      summarize(avgHPI = mean(CSUSHPISA),
                avgCPI = mean(CPIAUCSL))
```

## Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
#Use v_data from previous tests
hpi_cpi_plot <- plot_hpi_cpi(v_data); hpi_cpi_plot
```

## Correlation plot
```{r}
numeric_vars <- c()

for (i in colnames(viz_data)){
  
  if(i == "imputedYN"){
    next
    
  }else if (is.numeric(viz_data[,i]) == TRUE){
    
    numeric_vars <- append(numeric_vars, i)
  }
}

correlations <- cor(na.omit(viz_data[,c(numeric_vars)]))

corr_plot <- ggcorrplot(corr =  correlations,
                    colors = c(palette[1], palette[2], palette[3]),
                    ggtheme = "theme_bw", title = "Correlations", 
                    tl.cex = 8, legend.title = "") +
  theme(plot.title = element_text(face = "bold"))

corr_plot
```

## Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
#Use v_data from previous test
plot_correlations(v_data)
```

## Correlation tables
```{r}
library(GGally)
library(flextable)

correlations_df <- data.frame(correlations)

imp_vars <- correlations_df %>% 
  select(CSUSHPISA) %>% 
  filter(abs(CSUSHPISA) > 0.50)

imp_vars <- imp_vars %>%
  rename(hpi = CSUSHPISA) %>%
  mutate(name = rownames(imp_vars)) %>%
  relocate(name)
```

## Test visualization functions
```{r}
imp_varsss <- filter_imp_vars(v_data)
```

## Scatter plot matrix
```{r}
filtered_data <- viz_data[,c(unique(imp_vars$name))]

plots <- list()
ind = 1
hpi_idx <- which(colnames(filtered_data) == "CSUSHPISA")

for (i in 1:ncol(filtered_data)){
  plots[[ind]] <- ggplot(filtered_data, 
                         aes_string(x = filtered_data[,i], 
                                    y = filtered_data[,hpi_idx])) +
    geom_point(na.rm = TRUE) + 
    geom_smooth(na.rm = TRUE, se = FALSE, color = palette[3], 
                formula = y ~ x, method = "loess", alpha = 0.50) +
    labs(x = colnames(filtered_data)[i], y = "HPI") + 
    theme_bw()
  
    ind = ind + 1
}

grid.arrange(grobs = plots, nrow = 4, 
                   top = textGrob(
                     substitute(
                       paste(
                         bold("Variables with absolute correlation > 0.50 with HPI")))))
```

## Test visualization functions
```{r}
source("../scripts/visualization_functions.R")
plot_imp_line_matrix(viz_data = v_data, imp_vars = imp_varsss, hpi_col = "CSUSHPISA")
```








